Prestige as a Driving Force in Cultural Transmission

### 2022-03-28 ms/PrestigeBias.tex
revise name of alpha; add todos for fig 1 and 2

\usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amsmath, amssymb, amsthm}
\usepackage[font=small]{caption, subcaption}
\usepackage{setspace}\onehalfspacing
\usepackage[loose,nice]{units}
\usepackage{array}
\usepackage[super]{nth}
\usepackage{graphicx}
\usepackage{float}
\usepackage[unicode=true,colorlinks=true,urlcolor=Blue,citecolor=black,linkcolor=black]{hyperref}
\usepackage{cleveref}
\usepackage{subcaption}
\usepackage{mathtools}
\usepackage[all]{nowidow}
\usepackage{wrapfig}
\usepackage{pdfpages}
\usepackage{authblk}

@@

Following \cref{eq:ourF}, the probability that the $i$-th copier chose role-model $j$ is given by the prestige of role-model $j$ in the eyes of copier $i$,
\begin{equation}\label{eq:recPrestige}
Pr(S_{i,j}=1 \mid S_{1,j},S_{2,j},...,S_{i-1,j}) = G_{i,j} \;.
\end{equation}
The prestige $G_{i,j}$ of role-model~$j$ in the eyes of copier~$i$ is determined by the estimated biased indicator value $\beta(A_{i,j})$ and the number of copiers that chose role-model $j$ before copier $i$, $K_{i-1,j}$, replacing \cref{eq:boydG} of \citet{evolutionBook} with
\begin{equation}\label{eq:prestige}
G_{i,j} = \frac{\alpha_j \cdot \beta(A_{i,j}) + (1-\alpha_j) \cdot K_{i-1,j}}{W_i} \;.
\end{equation}
Here, the weight $\alpha_j$ is a characteristic of role-model~$j$ that determines the relative significance of success and influence in his prestige, the indicator value of role-model $j$ estimates by copier $i$, $A_{i,j}$, remains as in \cref{eq:relativeIndicator}, and $W_i$ is a normalizing factor that sums the numerator over all role-models ($1\le j \le N)$ to ensure $\sum_{j=1}^{N}{G_{i,j}}=1$.
%\begin{equation}
%W_i = \sum_{j=1}^N{\alpha_j \cdot \beta(A_{i,j}) + (1-\alpha_j) \cdot K_{i-1,j}}.
%\end{equation}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Results}
\subsection*{Approximations}
Our model is defined by two nested stochastic processes. Change over multiple generations is described by the phenotype distribution at each generation, $\{\vec{A}(t)\}_t$. The transition from one generation to the next is described by the number of copiers each role-model has after $i$ copiers have chosen a role-model, $\{\vec{K}_i\}_{i=1}^N$.
We emphasize that the models are nested: $\vec{A}(t+1)$ can be computed from $\vec{A}(t)$ by evaluating $\vec{K}_{N}$. However, the former requires iterating over \cref{eq:recurrence,eq:recPrestige}. Thus, we sought to find an equivalent stochastic process that has the same joint distribution as $\vec{K}_{N}$. 
We found two approximations such processes, summarized here and explained in detail below:
\begin{enumerate}
\item 
The number of copiers of a specific role-model at each step, $K_{i,j}$, follows the {\em generalized binomial distribution}~\citep{GBD} and therefore,
\begin{enumerate}\item the expected number of copiers of role-model~$j$ equals its prestige in the eyes of the first copier, multiplied by the total number of copiers, that is,
$\E[K_{N,j}] = N \cdot G_{1,j}$ if trait estimation error is uniform for all copiers ($e=e_i$ for $i=1, \ldots, N$).
\item the expected number of copiers of each role-model equals its relative biased indicator value, similar to the role of relative fitness in population-genetic models, that is, $\E[K_{N,j}] = \beta(A_j+e) / \overline{\beta}$ if the success weight is uniform for all role-models ($\alpha=\alpha_j$ for $j=1,\ldots,N$), where $\overline{\beta}=1/N \sum_{j=1}^{N}{\beta(A_j+e)}$ is the population mean estimated indicator value. 
\end{enumerate}
\item The role-model choice process, $\{\vec{K}_{i}\}_{i=1}^{N}$, is equivalent to a {\em P\'{o}lya urn} model if trait estimation error is uniform for all copiers ($e=e_i$ for all $i=1, \ldots, N$). 
Therefore, the number of copiers of all role-models, $\vec{K}_i = (K_{i,1}, \ldots, K_{i,N})$, follows a Dirichlet-Multinomial distribution,
$\vec{K}_i \sim \mathit{DM}\big(N, \vec{G}_1\big)$, where $\vec{G}_1 = (G_{1,1}, \ldots, G_{1,N})$ is the prestige vector of all role-models in the "eyes" of the first copier.
Note that here $G_{i,j}$ is only a function of the indicator values $A_j$ and the weights $\alpha_j$, as the estimation error is uniform. % TODO correct? anything else need to be assumed?
\end{enumerate}

@@

The special case where the weight is uniform for all role-models ($\alpha = \alpha_j$ for $j=1,\ldots, N$) is interesting, because we can evaluate the expected number of copiers using a linear equation
\begin{equation}
\E[K_{Nj}]= N\cdot\frac{\alpha\cdot\beta(A_j+e)}{\sum\limits_{m=1}^{N}\alpha\cdot\beta(A_m+e)} =\beta(A_j+e) \bigg/ \overline{\beta(A+e)} \;,
\end{equation}
where the only variable is $A_j+e$, because $\overline{\beta(A+e)}$ is the mean of the distribution we draw the indicator values from, modified by some constant parameters of $\beta$.
We can then denote $L = 1/\overline{\beta(A+e)}$ and write
\begin{equation}\label{eq:linearEq}
\E[K_{Nj}] = L\cdot \beta(A_j+e) \;.
\end{equation}

@@

\begin{result}[P{\'{o}}lya urn model]\label{result:polya}
The role-model choice process, $\big\{\vec{K}_i\big\}_{i=1}^N$, is equivalent to a \textit{P\'{o}lya urn model} if both trait estimation error and success weight are uniform in the population, $e=e_j$ and $\alpha=\alpha_j$ for all $j=1,\ldots,N$.
\end{result}

\begin{proof} 
Denote $\alpha'=\frac{\alpha}{1-\alpha}$ as the odds ratio between the weights of success and influence bias and $A'_j=A_j+e$. From \cref{eq:prestige} and because $\sum_{j=1}^{N}{K_{i,j}}=i$, we have
\begin{equation}\label{eq:copier_choose}
G_{i,j} = 
\frac{\alpha'\beta(A'_j) + K_{i-1,j}}{\sum\limits_{m=1}^{N} \alpha'\beta(A'_m) + K_{i-1,m}}
 =\frac{\alpha'\beta(A'_j) + K_{i-1,j}}{i-1 + \sum\limits_{m=1}^{N}\alpha'\beta(A'_m)} \;.
\end{equation}
Substituting $M=N$, $o_j = \alpha'\beta(A'_j)$, and $w_{i,j} = K_{i,j}$ in \cref{eq:polya} gives \cref{eq:copier_choose}, thus completing the proof.
\end{proof}

@@

\subsection*{Numeric comparisons} % TODO rename section
We're interesting in studying the difference between the real binary model as we defined in \cref{eq:binary-model}, and the Dirichlet-Multinomial approximation. Specifically, we're interesting in the fixation probability of the favored trait ($\hat{A}$) and its time to fixation.

\begin{figure}
    \includegraphics[width=\linewidth]{../figures/binary/num_sims.pdf}
  \caption{The number of simulations needed to get a good approximation. At $1,000$ the approximation is good enough. Error bars represent 95\% confidence interval. Population size $N=1000$, $\alpha=0.5$, $J=1$, $\hat{A}=1$,$A=0.7$, $\beta(A)=0.956$.}	
  \label{fig:num_sims}
\end{figure}

The first step was to find the number of simulations needed to sufficiently approximate the real model with the DM approximation. From \cref{fig:num_sims} we see that $1000$ simulations or higher is enough.

The next step was to see how the observed metrics (fixation probability and time) varies when relaxing our assumptions we used to prove the DM approximation.
First we relaxed our assumption of no mutation. To include mutation in the binary model, it needs to be redefined, since in the original model it was based on the fact the traits are drawn from a continuous scale. In the binary model mutation will be manifested as an error when evaluating the bias itself. This is easily done by using a heterogeneous $J$ parameter, which controls the strength of the success bias in \cref{eq:success_bias}.

\begin{figure}
    \includegraphics[width=\linewidth]{../figures/binary/full_vs_dm_mutation.pdf}
  \caption{Comparison of the DM approximation and the full model when mutation is included. Even high mutation rate doesn't worsen the approximation, and the data points are close to the mathematical estimation (Kimura's). Error bars are 95\% confidence intervals, and are condensed (+- 0.01 probability and +-40 generations)
  $5000$ simulations per data point, $N=1000$, $\alpha=0.1$,$\hat{A}=1$,$A=0.7$, $J\sim N(1,x^2)$ where $x$ is the mutation scale in the x-axis.
  }	
  \label{fig:hetro_mutation}
\end{figure}

In \cref{fig:hetro_mutation} we see the comparison when heterogeneous mutation is applied to both models. When mutation is applied, we sample $J_i$ for each copier $i$ from a normal distribution with varying scale (variance).
We can see that even when the standard deviation is $0.1$, the metrics of both models are both similar, and close to the Kimura approximation (more details in the next section).

@@

We again tried different variations of the parameters, changing only one of them at a time.
In \cref{fig:ch_env_alpha_beta} we can see that $\alpha$ on it's own doesn't cause any deviation for the the estimation. $\beta$ however affects the results greatly.

>>>>

\usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amsmath, amssymb, amsthm}
\usepackage[font=small,labelfont=bf]{caption, subcaption}
\usepackage{setspace}\onehalfspacing
\usepackage[loose,nice]{units}
\usepackage{array}
\usepackage[super]{nth}
\usepackage{graphicx}
\usepackage{float}
\usepackage[unicode=true,colorlinks=true,urlcolor=Blue,citecolor=black,linkcolor=black]{hyperref}
\usepackage{cleveref}
\usepackage{subcaption}
\usepackage{mathtools}
\usepackage[all]{nowidow}
\usepackage{wrapfig}
\usepackage{pdfpages}
\usepackage{authblk}

@@

Following \cref{eq:ourF}, the probability that the $i$-th copier chose role-model $j$ is given by the prestige of role-model $j$ in the eyes of copier $i$,
\begin{equation}\label{eq:recPrestige}
Pr(S_{i,j}=1 \mid S_{1,j},S_{2,j},...,S_{i-1,j}) = G_{i,j} \;.
\end{equation}
The prestige $G_{i,j}$ of role-model~$j$ in the eyes of copier~$i$ is determined by the estimated biased indicator value $\beta(A_{i,j})$ and the number of copiers that chose role-model $j$ before copier $i$, $K_{i-1,j}$, replacing \cref{eq:boydG} of \citet{evolutionBook} with
\begin{equation}\label{eq:prestige}
G_{i,j} = \frac{\alpha_j \cdot \beta(A_{i,j}) + (1-\alpha_j) \cdot K_{i-1,j}}{W_i} \;.
\end{equation}
Here, the bias weight $\alpha_j$ is a characteristic of role-model~$j$ that determines the relative significance of success and influence in his prestige, the indicator value of role-model $j$ estimates by copier $i$, $A_{i,j}$, remains as in \cref{eq:relativeIndicator}, and $W_i$ is a normalizing factor that sums the numerator over all role-models ($1\le j \le N)$ to ensure $\sum_{j=1}^{N}{G_{i,j}}=1$.
%\begin{equation}
%W_i = \sum_{j=1}^N{\alpha_j \cdot \beta(A_{i,j}) + (1-\alpha_j) \cdot K_{i-1,j}}.
%\end{equation}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Results}
\subsection*{Approximations}
Our model is defined by two nested stochastic processes. Change over multiple generations is described by the phenotype distribution at each generation, $\{\vec{A}(t)\}_t$. The transition from one generation to the next is described by the number of copiers each role-model has after $i$ copiers have chosen a role-model, $\{\vec{K}_i\}_{i=1}^N$.
We emphasize that the models are nested: $\vec{A}(t+1)$ can be computed from $\vec{A}(t)$ by evaluating $\vec{K}_{N}$. However, the former requires iterating over \cref{eq:recurrence,eq:recPrestige}. Thus, we sought to find an equivalent stochastic process that has the same joint distribution as $\vec{K}_{N}$. 
We found two approximations such processes, summarized here and explained in detail below:
\begin{enumerate}
\item 
The number of copiers of a specific role-model at each step, $K_{i,j}$, follows the {\em generalized binomial distribution}~\citep{GBD} and therefore,
\begin{enumerate}\item the expected number of copiers of role-model~$j$ equals its prestige in the eyes of the first copier, multiplied by the total number of copiers, that is,
$\E[K_{N,j}] = N \cdot G_{1,j}$ if trait estimation error is uniform for all copiers ($e=e_i$ for $i=1, \ldots, N$).
\item the expected number of copiers of each role-model equals its relative biased indicator value, similar to the role of relative fitness in population-genetic models, that is, $\E[K_{N,j}] = \beta(A_j+e) / \overline{\beta}$ if the bias weight is uniform for all role-models ($\alpha=\alpha_j$ for $j=1,\ldots,N$), where $\overline{\beta}=1/N \sum_{j=1}^{N}{\beta(A_j+e)}$ is the population mean estimated indicator value. 
\end{enumerate}
\item The role-model choice process, $\{\vec{K}_{i}\}_{i=1}^{N}$, is equivalent to a {\em P\'{o}lya urn} model if trait estimation error is uniform for all copiers ($e=e_i$ for all $i=1, \ldots, N$). 
Therefore, the number of copiers of all role-models, $\vec{K}_i = (K_{i,1}, \ldots, K_{i,N})$, follows a Dirichlet-Multinomial distribution,
$\vec{K}_i \sim \mathit{DM}\big(N, \vec{G}_1\big)$, where $\vec{G}_1 = (G_{1,1}, \ldots, G_{1,N})$ is the prestige vector of all role-models in the "eyes" of the first copier.
Note that here $G_{i,j}$ is only a function of the indicator values $A_j$ and the bias weights $\alpha_j$, as the estimation error is uniform. % TODO correct? anything else need to be assumed?
\end{enumerate}

@@

The special case where the bias weight is uniform for all role-models ($\alpha = \alpha_j$ for $j=1,\ldots, N$) is interesting, because we can evaluate the expected number of copiers using a linear equation
\begin{equation}
\E[K_{Nj}]= N\cdot\frac{\alpha\cdot\beta(A_j+e)}{\sum\limits_{m=1}^{N}\alpha\cdot\beta(A_m+e)} =\beta(A_j+e) \bigg/ \overline{\beta(A+e)} \;,
\end{equation}
where the only variable is $A_j+e$, because $\overline{\beta(A+e)}$ is the mean of the distribution we draw the indicator values from, modified by some constant parameters of $\beta$.
We can then denote $L = 1/\overline{\beta(A+e)}$ and write
\begin{equation}\label{eq:linearEq}
\E[K_{Nj}] = L\cdot \beta(A_j+e) \;.
\end{equation}

@@

\begin{result}[P{\'{o}}lya urn model]\label{result:polya}
The role-model choice process, $\big\{\vec{K}_i\big\}_{i=1}^N$, is equivalent to a \textit{P\'{o}lya urn model} if both trait estimation error and bias weight are uniform in the population, $e=e_j$ and $\alpha=\alpha_j$ for all $j=1,\ldots,N$.
\end{result}

\begin{proof} 
Denote $\alpha'=\frac{\alpha}{1-\alpha}$ as the bias weight ratio, and $A'_j=A_j+e$. From \cref{eq:prestige} and because $\sum_{j=1}^{N}{K_{i,j}}=i$, we have
\begin{equation}\label{eq:copier_choose}
G_{i,j} = 
\frac{\alpha'\beta(A'_j) + K_{i-1,j}}{\sum\limits_{m=1}^{N} \alpha'\beta(A'_m) + K_{i-1,m}}
 =\frac{\alpha'\beta(A'_j) + K_{i-1,j}}{i-1 + \sum\limits_{m=1}^{N}\alpha'\beta(A'_m)} \;.
\end{equation}
Substituting $M=N$, $o_j = \alpha'\beta(A'_j)$, and $w_{i,j} = K_{i,j}$ in \cref{eq:polya} gives \cref{eq:copier_choose}, thus completing the proof.
\end{proof}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[h]
    \includegraphics[width=\linewidth]{../figures/binary/num_sims.pdf}
  \caption{
  \textbf{Number of simulations needed to get a good approximation.}
  The approximation (orange) fits simulation results (blue) well when using 1,000 simulations. % 1000 simulations of the model or of DM ??
  Markers for average value across simulations, error bars for 95\% confidence interval.
  Here, population size, $N=1000$; bias weight, $\alpha=0.5$; phenotype values, $\hat{A}=1$, $A=0.7$; success bias value, $\beta(A)=0.956$.}	
  \label{fig:num_sims}
  % TODO
  % remove legend
  % change y labels: "Fixation probability" "Fixation time"
  % change x label: "Number of simulations"
  % Kimura - black solid line
  % I removed J as you give beta(A) below, ok
  % it is not clear what "number of simulations needed" means.
  % what does "good enough" mean?
  % maybe y axis should be in relative error?
  % you should point to the equation for kimura.
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection*{Numeric comparisons} % TODO rename section

Here we examine the difference between the dichotomous trait model defined in \cref{eq:binary-model} and the Dirichlet-Multinomial approximation we presented in \cref{cor:dirichlet}. We focus on the fixation probability and fixation time of the advantageous phenotype $\hat{A}$ when invading a population of phenotype $A$.

First, we find that the number of simulations needed to sufficiently approximate our model with the DM approximation is roughly $1000$ (\Cref{fig:num_sims}).
Next, we examined the robustness of the DM approximation to relaxing the approximation assumptions
First we relaxed our assumption of no mutation. % TODO what is the mutation parameters? e? I thought this was estimation error. does it only affect the bias or also the copying? we should discuss!
To include mutation in the dichotomous trait model, it needs to be redefined, since in the original model it was based on the fact the traits are drawn from a continuous scale. In the dichotomous trait model mutation will be manifested as an error when evaluating the bias itself. This is easily done by using a heterogeneous $J$ parameter, which controls the strength of the success bias in \cref{eq:success_bias}. % NOT CLEAR AT ALL - do you redefine now the model? or is this leftover from before??? 
% I dont like that we call this mutation, we should call it copying error or something else, see if Boyd or Fogarty has a name for it.

\begin{figure}
    \includegraphics[width=\linewidth]{../figures/binary/full_vs_dm_mutation.pdf}
  \caption{
  \textbf{Robustness of DM approximations to inclusion of mutation.} Both the DM approximation (orange) and Kimura's equation (black line) fit the stochastic simulations (blue) well even with a high mutation rate. Markers for average across simulations, error bars for 95\% confidence intervals.
  % and are condensed (+- 0.01 probability and +-40 generations) % what does that mean?
  5,000 simulations per data point; population size, $N=1000$; bias weight, $\alpha=0.1$; phenotype values, $\hat{A}=1$,$A=0.7$; $J\sim N(1,x^2)$ where $x$ is the mutation scale in the x-axis. % not clear enough.
  }	
  % TODO
  % see comments in previous figure
  \label{fig:hetro_mutation}
\end{figure}

% YOAV HERE
In \cref{fig:hetro_mutation} we see the comparison when heterogeneous mutation is applied to both models. When mutation is applied, we sample $J_i$ for each copier $i$ from a normal distribution with varying scale (variance).
We can see that even when the standard deviation is $0.1$, the metrics of both models are both similar, and close to the Kimura approximation (more details in the next section).

@@

We again tried different variations of the parameters, changing only one of them at a time.
In \cref{fig:ch_env_alpha_beta} we can see that $\alpha$ on its own does not cause any deviation for the the estimation. $\beta$ however affects the results greatly.
